---
title: "Difference in BF computed by numerial intergration vs. ABF"
author: "Yunqi Yang"
date: "7/31/2023"
output: html_document
---

### Description: 

Assess difference in Bayes factor computed by numerical integration vs. ABF. The example of using numerical integration to compute BF is available at: https://yunqiyang0215.github.io/survival-susie/numerical_integration.html


We assess three scenarios under relatively large sample size, $n= 10,000$.

1. Extremely high censoring rate, censoring rate = 0.99

2. Low allele frequency, MAF = 0.001

3. Both. 


Result shows that low allele frequency makes the difference between BF and ABF larger. 


```{r}
# Function to simulate survival time under exponential model. That is,
# assuming survival time is exponentially distributed.
# lambda(t) = lambda*exp(b0 + Xb). S(t) = exp(-lambda*t),
# F(t) = 1- S(t) \sim Unif(0,1). Therefore, t = log(1-runif(0,1))/-exp(b0+Xb).
# For censored objects, we simulate the censoring time by rescale the actual survival time.
# @param b: vector of length (p+1) for true effect size, include intercept.
# @param X: variable matrix of size n by p.
# @param censor_lvl: a constant from [0,1], indicating the censoring level in the data.
# @return  dat: a dataframe that contains `y`, `x` and `status`.
# `status`: censoring status: 0 = censored, 1 = event observed. See Surv() in library(survival)
sim_surv_exp <- function(b, X, censor_lvl){
  n = nrow(X)
  p = ncol(X)
  dat = list()

  status <- ifelse(runif(n) > censor_lvl, 1, 0)
  lambda <- exp(cbind(rep(1,n), X) %*% b)
  surT <- log(1 - runif(n)) /(-lambda)
  # rescale censored subject to get observed time
  surT[status == 0] = surT[status == 0] * runif(sum(status == 0))

  y = cbind(surT, status)
  colnames(y) = c("time", "status")
  colnames(X) <- unlist(lapply(1:p, function(i) paste0("x", i)))
  dat[["X"]] = X
  dat[["y"]] = y
  return(dat)
}

compute_multiple_lbf <- function(x, y, o, prior_variance, zscore.spa, ...){
  fit <- coxph(y~ x + offset(o))
  bhat <- summary(fit)$coefficients[1, 1] # bhat = -alphahat
  sd <- summary(fit)$coefficients[1, 3]
  zscore <- bhat/sd
  lbf <- compute_lbf(zscore, sd, prior_variance)
  lbf.corr <- lbf - bhat^2/sd^2/2+ as.numeric(summary(fit)$logtest[1]/2)
  lbf.spa <- compute_lbf(zscore.spa, sd, prior_variance)
  return(list(zscore = zscore, sd = sd, 
              lbf=lbf, lbf.corr = lbf.corr, lbf.spa = lbf.spa))
}
```

```{r}
# @param b: the effect size
# @param surT: a Surv() object, containing time and status
# @param x: covariate
get_partial_lik <- function(b, survT, x){
  ll.partial = logLik(coxph(survT ~ x, init = c(b), control=list('iter.max'= 0, timefix=FALSE)))
  max.ll.partial = logLik(coxph(survT ~ x))
  lik.partial.normed = exp(ll.partial - max.ll.partial) # partial likelihood / max(partial likelihood)
  return(as.numeric(lik.partial.normed))
}

integrand <- function(b, survT, x, prior_sd){
  val = get_partial_lik(b, survT, x) * dnorm(b, mean = 0, sd = prior_sd)
  return(val)
}
```

```{r}
get_zscore_spa <- function(dat, X){
  # genome matrix
  n = nrow(X)
  Geno.mtx = X
  Phen.mtx = data.frame(ID = paste0("IID-",1:n),
                      event=dat$y[,2],
                      time=dat$y[,1])
  rownames(Geno.mtx) = paste0("IID-",1:nrow(Geno.mtx))
  colnames(Geno.mtx) = paste0("SNP-",1:ncol(Geno.mtx))
  obj.null = SPACox_Null_Model(Surv(time,event) ~ 1, data=Phen.mtx, 
                               pIDs=Phen.mtx$ID, gIDs=rownames(Geno.mtx))
  SPACox.res = SPACox(obj.null, Geno.mtx)
  zscore = SPACox.res[1, 7]
  return(zscore)
}
```

```{r}
library(survival)
library(cubature)
library(SPACox)
source("./code/surv_susie_helper.R")

```


### Scenario 1:
```{r}
seeds = c(1:5)
result = matrix(NA, ncol = 5, nrow = length(seeds))
colnames(result) = c("lbf.wakefeld", "lbf.laplace", "lbf.numerical", "lbf.spa", "error")
for (seed in seeds){
  set.seed(seed)
  # simulate 1 variable
  n = 1e4
  X = cbind(rbinom(n, size = 2, prob = 0.2))
  # the first element of b is for intercept
  b = c(1, 1)
  censor_lvl = 0.99
  dat <- sim_surv_exp(b, X, censor_lvl)
  survT <- Surv(dat$y[,1], dat$y[,2])

  # compute lbf using numerical integration 
  prior_sd = 1
  res = cubintegrate(f = integrand, survT = survT, x = dat$X[,1], prior_sd = 1, lower = -100, upper = 100, method = "hcubature")
  max.ll.partial = logLik(coxph(survT ~ dat$X[,1]))
  prob.h1 = res$integral*exp(max.ll.partial)
  prob.h0 = get_partial_lik(b = 0, survT, dat$X[,1])
  lbf.numerical = log(res$integral) - log(prob.h0)
  
  # compute other lbf
  zscore.spa <- get_zscore_spa(dat, X)
  res.lbf = compute_multiple_lbf(x = dat$X[,1], y = survT, o = rep(0, n), prior_variance = prior_sd^2, zscore.spa = zscore.spa)
  
  
  result[seed, 1] = res.lbf$lbf
  result[seed, 2] = res.lbf$lbf.corr
  result[seed, 3] = lbf.numerical
  result[seed, 4] = res.lbf$lbf.spa
  result[seed, 5] = res$error
}

```

```{r}
result
```

### Scenario 2:
```{r}
seeds = c(1:5)
result = matrix(NA, ncol = 5, nrow = length(seeds))
colnames(result) = c("lbf.wakefeld", "lbf.laplace", "lbf.numerical", "lbf.spa", "error")
for (seed in seeds){
  set.seed(seed)
  # simulate 1 variable
  n = 1e4
  X = cbind(rbinom(n, size = 2, prob = 0.001))
  # the first element of b is for intercept
  b = c(1, 1)
  censor_lvl = 0.6
  dat <- sim_surv_exp(b, X, censor_lvl)
  survT <- Surv(dat$y[,1], dat$y[,2])

  
  # compute lbf using numerical integration 
  prior_sd = 1
  res = cubintegrate(f = integrand, survT = survT, x = dat$X[,1], prior_sd = 1, lower = -100, upper = 100, method = "hcubature")
  max.ll.partial = logLik(coxph(survT ~ dat$X[,1]))
  prob.h1 = res$integral*exp(max.ll.partial)
  prob.h0 = get_partial_lik(b = 0, survT, dat$X[,1])
  lbf.numerical = log(res$integral) - log(prob.h0)
  
  # compute other lbf
  zscore.spa <- get_zscore_spa(dat, X)
  res.lbf = compute_multiple_lbf(x = dat$X[,1], y = survT, o = rep(0, n), prior_variance = prior_sd^2, zscore.spa = zscore.spa)
  
  
  result[seed, 1] = res.lbf$lbf
  result[seed, 2] = res.lbf$lbf.corr
  result[seed, 3] = lbf.numerical
  result[seed, 4] = res.lbf$lbf.spa
  result[seed, 5] = res$error
}

```

```{r}
result
```


### Scenario 3:
```{r warning=FALSE}
seeds = c(1:3)
result = matrix(NA, ncol = 5, nrow = length(seeds))
colnames(result) = c("lbf.wakefeld", "lbf.laplace", "lbf.numerical", "lbf.spa", "error")
for (seed in seeds){
  set.seed(seed)
  # simulate 1 variable
  n = 1e4
  X = cbind(rbinom(n, size = 2, prob = 0.001))
  # the first element of b is for intercept
  b = c(1, 1)
  censor_lvl = 0.99
  dat <- sim_surv_exp(b, X, censor_lvl)
  survT <- Surv(dat$y[,1], dat$y[,2])

  
  # compute lbf using numerical integration 
  prior_sd = 1
  res = cubintegrate(f = integrand, survT = survT, x = dat$X[,1], prior_sd = 1, lower = -100, upper = 100, method = "hcubature")
  max.ll.partial = logLik(coxph(survT ~ dat$X[,1]))
  prob.h1 = res$integral*exp(max.ll.partial)
  prob.h0 = get_partial_lik(b = 0, survT, dat$X[,1])
  lbf.numerical = log(res$integral) - log(prob.h0)
  
  # compute other lbf
  zscore.spa <- get_zscore_spa(dat, X)
  res.lbf = compute_multiple_lbf(x = dat$X[,1], y = survT, o = rep(0, n), prior_variance = prior_sd^2, zscore.spa = zscore.spa)
  
  
  result[seed, 1] = res.lbf$lbf
  result[seed, 2] = res.lbf$lbf.corr
  result[seed, 3] = lbf.numerical
  result[seed, 4] = res.lbf$lbf.spa
  result[seed, 5] = res$error
}

```

```{r}
result
```

