---
title: "bhglm"
author: "Yunqi Yang"
date: "3/6/2023"
output: html_document
---

## Description:

The candidate method is called BhGLM. 

Publication link: https://academic.oup.com/bioinformatics/article/33/18/2799/3796395?login=true 

Software link: https://github.com/nyiuab/BhGLM/blob/master/vignettes/Cox.pdf

```{r knitr, echo=FALSE}
knitr::opts_chunk$set(comment = "#",results = "hold",collapse = TRUE,
                      fig.align = "center")
```

```{r}
library(survival)
library(BhGLM)
```

```{r}
dat = readRDS("./data/sim_dat_simple.rds")
```

### Data 1: null model with X independent
```{r}
# Create survival response
dat[[1]]$y <- with(dat[[1]], Surv(surT, status == 2))

p = 50
X = as.matrix(dat[[1]][, c(2:(p+1))])
y = dat[[1]]$y

```

```{r}
fit <- bcoxph(y ~ X, prior = mde(0, s0 = 0.05, s1 = 0.5))
out <- summary.bh(fit)
out
```


### Data 2: simulated from null model with highly correlated X. corr = 0.9
```{r}
# Create survival response
dat[[2]]$y <- with(dat[[2]], Surv(surT, status == 2))

p = 50
X = as.matrix(dat[[2]][, c(2:(p+1))])
y = dat[[2]]$y
```

```{r}
fit <- bcoxph(y ~ X, prior = mde(0, s0 = 0.05, s1 = 0.5))
out <- summary.bh(fit)
out
```

### Data 3: simulated from one predictor model. Predictors are independent.

```{r}
# Create survival response
dat[[3]]$y <- with(dat[[3]], Surv(surT, status == 2))

p = 50
X = as.matrix(dat[[3]][, c(2:(p+1))])
y = dat[[3]]$y
```

```{r}
fit <- bcoxph(y ~ X, prior = mde(0, s0 = 0.05, s1 = 0.5))
out <- summary.bh(fit)
out
```

### Data 4: simulated from one predictor model. Predictors are highly correlated, corr = 0.9
```{r}
# Create survival response
dat[[4]]$y <- with(dat[[4]], Surv(surT, status == 2))

p = 50
X = as.matrix(dat[[4]][, c(2:(p+1))])
y = dat[[4]]$y
```

```{r}
fit <- bcoxph(y ~ X, prior = mde(0, s0 = 0.05, s1 = 0.5))
out <- summary.bh(fit)
out
```

### Data 5: simulated from two predictor model. Predictors have high correlation, corr = 0.9
```{r}
# Create survival response
dat[[5]]$y <- with(dat[[5]], Surv(surT, status == 2))

p = 50
X = as.matrix(dat[[5]][, c(2:(p+1))])
y = dat[[5]]$y
```

```{r}
fit <- bcoxph(y ~ X, prior = mde(0, s0 = 0.05, s1 = 0.5))
out <- summary.bh(fit)
out
```
