---
title: "bvsnlp with censoring"
author: "Yunqi Yang"
date: "3/15/2023"
output: html_document
---

## Description:

Run **bvsnlp** on censoring data. 

Publication link: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7808442/

Software link: https://cran.r-project.org/web/packages/BVSNLP/index.html

**Conclusion**: This method doesn't perform well under the null model, picked one or two variables under the null model. For other 3 scenarios, it performs well.

```{r}
library(survival)
library(BVSNLP)
```

```{r}
dat = readRDS("./data/sim_dat_censoring.rds")
```


## Main function:
```{r eval = FALSE}
bvs(
  X,
  resp,
  prep = TRUE, # If preprocessing 
  logT = FALSE,
  fixed_cols = NULL,
  eff_size = 0.5, # expected effect size
  family = c("logistic", "survival"),
  hselect = TRUE, # whether the automatic procedure for hyperparameter selection should be run or not.
  nlptype = "piMOM",
  r = 1,
  tau = 0.25,
  niter = 30,
  mod_prior = c("unif", "beta"),
  inseed = NULL,
  cplng = FALSE, # This parameter is only used in logistic regression models
  ncpu = 4,
  parallel.MPI = FALSE
)


```

### Data 1: null model with X independent
```{r}
# Create survival response
dat[[1]]$y <- with(dat[[1]], Surv(surT, status))

p = 50
X = as.data.frame(dat[[1]][, c(2:(p+1))])
y = dat[[1]]$y
```

```{r}
# prep: standardizing non-binary columns and adding intercept
fit1 = bvs(X, resp = y, prep = TRUE, family = "survival", mod_prior = "beta", niter = 100)
```

```{r}
# The coefficient vector for the selected model. 
# The first component is always for the intercept.
fit1$beta_hat
# The indices of the model with highest posterior probability among all visited
# models, with respect to the columns in the output des_mat.
fit1$HPM
hist(fit1$inc_probs, breaks = 30)
```

### Data 2: simulated from null model with highly correlated X. corr = 0.9
```{r}
# Create survival response
dat[[2]]$y <- with(dat[[2]], Surv(surT, status))
p = 50
X = as.data.frame(dat[[2]][, c(2:(p+1))])
y = dat[[2]]$y
```

```{r}
fit2 = bvs(X, resp = y, prep = TRUE, family = "survival", mod_prior = "beta", niter = 100)
```

```{r}
fit2$beta_hat
fit2$HPM
hist(fit2$inc_probs, breaks = 30)
```

### Data 3: simulated from one predictor model. Predictors are independent.
```{r}
# Create survival response
dat[[3]]$y <- with(dat[[3]], Surv(surT, status))

p = 50
X = as.data.frame(dat[[3]][, c(2:(p+1))])
y = dat[[3]]$y
```

```{r}
fit3 = bvs(X, resp = y, prep = TRUE, family = "survival", mod_prior = "beta", niter = 100)
```

```{r}
fit3$beta_hat
fit3$HPM
hist(fit3$inc_probs, breaks = 30)
```

### Data 4: simulated from one predictor model. Predictors are highly correlated, corr = 0.9

```{r}
# Create survival response
dat[[4]]$y <- with(dat[[4]], Surv(surT, status))

p = 50
X = as.data.frame(dat[[4]][, c(2:(p+1))])
y = dat[[4]]$y
```

```{r}
fit4 = bvs(X, resp = y, prep = TRUE, family = "survival", mod_prior = "beta", niter = 100)
```

```{r}
fit4$beta_hat
fit4$HPM
hist(fit4$inc_probs, breaks = 30)
```


### Data 5: simulated from two predictor model. Predictors have high correlation, corr = 0.9

```{r}
# Create survival response
dat[[5]]$y <- with(dat[[5]], Surv(surT, status))

p = 50
X = as.data.frame(dat[[5]][, c(2:(p+1))])
y = dat[[5]]$y
```

```{r}
fit5 = bvs(X, resp = y, prep = TRUE, family = "survival", mod_prior = "beta", niter = 100)
```

```{r}
fit5$beta_hat
fit5$HPM
hist(fit5$inc_probs, breaks = 30)
```
