---
title: "standardize_effect_coxph"
author: "Yunqi Yang"
date: "2/13/2023"
output: html_document
---

### Description: 

check if standardizing will change the coefficient of variables in coxph model. Let's still use the exponential model from https://yunqiyang0215.github.io/survival-susie/sim_survival.html

The exponential model models the mean of exponential as:
$$
\begin{split}
E(T_i) &= \mu_i\\
\log\mu_i &= \beta_0 + X\beta
\end{split}
$$
The hazard in exponential model is $\lambda_i = 1/\mu_i=\frac{1}{\exp\{\beta_0+X\beta\}}=\lambda_0\exp\{X\alpha\}$. Therefore, $\alpha$ is the output from coxph model and $\alpha=-\beta$. 

*Conclusion*: 

1. Operations only on x:

(1) Mean center won't change the coefficient in coxph model.

(2) The coefficient will be multiplied by $s$ if the variable is divided by $s$.


2. Standardize both x and y: the beta estimate is the same as the original data without standardize x and y. 

```{r}
library(mvtnorm)
library(survival)
```

```{r}
# Here we use parametric model to simulate data with survival time,
# assuming survival time is exponentially distributed. 
# We first simulate the mean of exponential from linear combinations
# of variables, and then simulate survival time. 
# T\sim 1/u*exp(-t/u), and the true model is:
# log(T) = \mu + e = b0 + Xb + e
# @param b: vector of length (p+1) for true effect size, include intercept.
# @param X: variable matrix of size n by p. 
# @param status: censoring status. 1 = censored, 2 = event observed. 
sim_dat <- function(b, X){
  n = nrow(X)
  p = ncol(X)
  mu <- exp(cbind(rep(1,n), X) %*% b)
  surT <- rexp(n, rate = 1/mu)
  dat <- data.frame(cbind(surT, X))
  x.name <- unlist(lapply(1:p, function(i) paste0("x", i)))
  names(dat) = c("surT", x.name)
  dat$status <- rep(2, n)
  return(dat)
}

```


```{r}
set.seed(1)
n = 200
b = c(1,0.1)
X = as.matrix(rnorm(n, mean = 3, sd = 2))
dat = sim_dat(b, X)

x.c = scale(X, center = TRUE, scale = FALSE)
x.sd = scale(X, center = TRUE, scale = TRUE)
```

```{r}
hist(dat$surT, breaks = 20)
```

### Effect of mean center/standardize X
```{r}
## Create  survival object. status == 2 is death
dat$y <- with(dat, Surv(surT, status == 2))
# Fit cox ph. Cox ph model with select multiple significant predictors..
cox1 <- coxph(y ~ x1, data =  dat)
cox2 <- coxph(dat$y ~ x.sd)
cox3 <- coxph(dat$y ~ x.c)
```


```{r}
coef(cox1)
coef(cox2)
coef(cox3)
```

```{r}
summary(cox1)
summary(cox2)
summary(cox3)
```

### Effect of standardize both y and X

```{r}
## Create  survival object. status == 2 is death
surT = scale(dat$surT, center = TRUE, scale = TRUE)
status = rep(2, length(surT))
dat$y <- Surv(surT, status)

# Fit cox ph.
cox4 <- coxph(y ~ x.sd, data =  dat)
summary(cox4)
```




