---
title: "Power vs. fdr: version 4"
author: "Yunqi Yang"
date: "01/29/2024"
output: html_document
---

### Description:

Comparing power vs. FDR across 4 methods. I vary the threshold for claiming effect variables based on marginal PIP value. 

### Data details:

1. $n=50000$ and $p=1000$.

2. SNPs with minor allele frequency < 1% were removed. 

### Fitting details:

1. SuSIE with laplace BF. 

2. SuSIE, survival svb runs for 10 iterations. The other 2 methods with 30 iterations. 


```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center")
```

```{r}

calculate_tpr_vs_fdr <- function(pip, is_effect, ts){
  res <- matrix(NA, nrow = length(ts), ncol = 2)
  colnames(res) = c("tpr", "fdr")
  for (i in 1:length(ts)){
    pred_pos = pip >= ts[i]
    tp = pip >= ts[i] & is_effect == 1
    fp = pip >= ts[i] & is_effect == 0
    tpr = sum(tp)/sum(is_effect)
    fdr = sum(fp)/sum(pred_pos)
    res[i, ] = c(tpr, fdr)
  }
  return(res)
}

# coverage: the proportion of CSs that contain an effect variable
# @param dat_indx: the indx for the data from dsc
calculate_cs_coverage = function(dat_indx){
  contain_status = c()
  for (indx in dat_indx){
     cs = susie$susie.cs[[indx]]$cs
     true_effect = which(susie$simulate.is_effect[[indx]] == 1)
     if (!is.null(cs)){
       for (j in 1:length(cs)){
         res = ifelse(sum(true_effect %in% unlist(cs[j])) ==  1, 1, 0)
         contain_status = c(contain_status, res)
        }
     }
  }
  coverage = sum(contain_status)/length(contain_status)
  return(coverage)
}


# @param dat_indx: the indx for the data from dsc
# @p: number of variables in each simulation replicate. 
get_cs_effect = function(dat_indx, p){
  cs_effect = c()
  for (indx in dat_indx){
    effect = rep(0, p)
    cs_effect_indx = c(unlist(susie$susie.cs[[indx]]$cs))
    effect[cs_effect_indx] = 1
    cs_effect = c(cs_effect, effect)
  }
  return(cs_effect)
}
```

```{r}

susie = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/susie.rds")
survsvb = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/survsvb.rds")
bvsnlp = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/bvsnlp.rds")
r2b = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/r2b.rds")
```

```{r}
t95 = 951
```

### 1. Results using real correlation structure from data

```{r, fig.width=12, fig.height=8}
par(mfrow = c(2,3), cex.axis = 1.5)
censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "real" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$svb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  pip.r2b = unlist(lapply(indx, function(x) r2b$r2b.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 1, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  res.r2b = calculate_tpr_vs_fdr(pip.r2b, is_effect, ts)
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0,1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Real correlation, effect 0-3", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  lines(res.r2b[,2], res.r2b[,1], type = "l", col = 4)
  
  points(res.susie[t95,2], res.susie[t95, 1])
  points(res.svb[t95,2], res.svb[t95, 1])
  points(res.bvsnlp[t95,2], res.bvsnlp[t95, 1])
  points(res.r2b[t95,2], res.r2b[t95, 1])
  
  legend("topleft", legend = c("susie", "survival.svb", "bvsnlp", "r2b"), col = c(1,2,3, 4), lty = 1)
}

```
The dots indicate PIP threshold = 0.95


### 2. Results using independent X, without data from null model. 

```{r, fig.width=12, fig.height=8}
par(mfrow = c(2,3),cex.axis = 1.5)
censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)

for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i] & susie$simulate.num_effect != 0)
  
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$svb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  pip.r2b = unlist(lapply(indx, function(x) r2b$r2b.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 1, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  res.r2b = calculate_tpr_vs_fdr(pip.r2b, is_effect, ts)
  
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 0.1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("X independent, effect 1-3", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  lines(res.r2b[,2], res.r2b[,1], type = "l", col = 4)
  
  points(res.susie[t95,2], res.susie[t95, 1])
  points(res.svb[t95,2], res.svb[t95, 1])
  points(res.bvsnlp[t95,2], res.bvsnlp[t95, 1])
  legend("bottomright", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```

The dots indicate PIP threshold = 0.95.


### 3. Results using independent X, with data from null model. 

```{r, fig.width=10, fig.height=8}
par(mfrow = c(2,3),cex.axis = 1.5)

censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)

for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$svb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  pip.r2b = unlist(lapply(indx, function(x) r2b$r2b.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 1, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  res.r2b = calculate_tpr_vs_fdr(pip.r2b, is_effect, ts)
  
  
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("X independent, effect 0-3", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  lines(res.r2b[,2], res.r2b[,1], type = "l", col = 4)
  
  points(res.susie[t95,2], res.susie[t95, 1])
  points(res.svb[t95,2], res.svb[t95, 1])
  points(res.bvsnlp[t95,2], res.bvsnlp[t95, 1])
  legend("bottomright", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```

The dots indicate PIP threshold = 0.95.


