---
title: "Power vs. fdr: version 4"
author: "Yunqi Yang"
date: "05/08/2023"
output: html_document
---

### Description:

Version 4 simulation results, comparing power vs. FDR across 3 methods. I vary the threshold for claiming effect variables based on marginal PIP value. 

**Difference between v3 and v4 simulation**:

1. Only susie has been changed, not the other two methods. 

2. Used corrected ABF instead of original Wakefeld ABF.

3. Computed susie credible sets. 

**Conclusion**:

1. After using corrected ABF, susie performance is improved. Now it's better than bvsnlp almost all the time. 

2. Power of the credible sets and coverage of credible sets also look fine. 

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center")
```

```{r}

calculate_tpr_vs_fdr <- function(pip, is_effect, ts){
  res <- matrix(NA, nrow = length(ts), ncol = 2)
  colnames(res) = c("tpr", "fdr")
  for (i in 1:length(ts)){
    pred_pos = pip >= ts[i]
    tp = pip >= ts[i] & is_effect == 1
    fp = pip >= ts[i] & is_effect == 0
    tpr = sum(tp)/sum(is_effect)
    fdr = sum(fp)/sum(pred_pos)
    res[i, ] = c(tpr, fdr)
  }
  return(res)
}

# coverage: the proportion of CSs that contain an effect variable
# @param dat_indx: the indx for the data from dsc
calculate_cs_coverage = function(dat_indx){
  contain_status = c()
  for (indx in dat_indx){
     cs = susie$susie.cs[[indx]]$cs
     true_effect = which(susie$simulate.is_effect[[indx]] == 1)
     if (!is.null(cs)){
       for (j in 1:length(cs)){
         res = ifelse(sum(true_effect %in% unlist(cs[j])) ==  1, 1, 0)
         contain_status = c(contain_status, res)
        }
     }
  }
  coverage = sum(contain_status)/length(contain_status)
  return(coverage)
}


# @param dat_indx: the indx for the data from dsc
# @p: number of variables in each simulation replicate. 
get_cs_effect = function(dat_indx, p){
  cs_effect = c()
  for (indx in dat_indx){
    effect = rep(0, p)
    cs_effect_indx = c(unlist(susie$susie.cs[[indx]]$cs))
    effect[cs_effect_indx] = 1
    cs_effect = c(cs_effect, effect)
  }
  return(cs_effect)
}
```

```{r}

susie = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/susie.rds")
susie.rss = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/susie_rss.rds")
susie.rss.varY = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/susie_rss_varY.rds")
survsvb = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/survsvb.rds")
bvsnlp = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/bvsnlp.rds")
r2b = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/r2b.rds")


```

### 1. Results using real correlation structure from data

```{r, fig.width=12, fig.height=8}
par(mfrow = c(2,3), cex.axis = 1.5)
censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "real" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$svb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  pip.r2b = unlist(lapply(indx, function(x) r2b$r2b.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 1, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  res.r2b = calculate_tpr_vs_fdr(pip.r2b, is_effect, ts)
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0,1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Real correlation, effect 0-3", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  lines(res.r2b[,2], res.r2b[,1], type = "l", col = 4)
  
  points(res.susie[96,2], res.susie[96, 1])
  points(res.svb[96,2], res.svb[96, 1])
  points(res.bvsnlp[96,2], res.bvsnlp[96, 1])
  points(res.r2b[96,2], res.r2b[96, 1])
  
  legend("topleft", legend = c("susie", "survival.svb", "bvsnlp", "r2b"), col = c(1,2,3, 4), lty = 1)
}

```
The dots indicate PIP threshold = 0.95


### 2. Results using independent X, without data from null model. 

```{r, fig.width=12, fig.height=8}
par(mfrow = c(2,3),cex.axis = 1.5)
censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)

for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i] & susie$simulate.num_effect != 0)
  
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$svb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  pip.r2b = unlist(lapply(indx, function(x) r2b$r2b.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 1, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  res.r2b = calculate_tpr_vs_fdr(pip.r2b, is_effect, ts)
  
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 0.1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("X independent, effect 1-3", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  lines(res.r2b[,2], res.r2b[,1], type = "l", col = 4)
  
  points(res.susie[96,2], res.susie[96, 1])
  points(res.svb[96,2], res.svb[96, 1])
  points(res.bvsnlp[96,2], res.bvsnlp[96, 1])
  legend("bottomright", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```

The dots indicate PIP threshold = 0.95.


### 3. Results using independent X, with data from null model. 

```{r, fig.width=10, fig.height=8}
par(mfrow = c(2,3),cex.axis = 1.5)

censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)

for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$svb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  pip.r2b = unlist(lapply(indx, function(x) r2b$r2b.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 1, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  res.r2b = calculate_tpr_vs_fdr(pip.r2b, is_effect, ts)
  
  
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("X independent, effect 0-3", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  lines(res.r2b[,2], res.r2b[,1], type = "l", col = 4)
  
  points(res.susie[96,2], res.susie[96, 1])
  points(res.svb[96,2], res.svb[96, 1])
  points(res.bvsnlp[96,2], res.bvsnlp[96, 1])
  legend("bottomright", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```

The dots indicate PIP threshold = 0.95.


