---
title: "Compare Power vs. FDR"
author: "Yunqi Yang"
date: "4/7/2023"
output: html_document
---

### Description:

Compare the power vs. FDR across 3 methods. I vary the threshold for claiming effect variables based on marginal PIP value. 

**Conclustion**:

1. When X are highly correlated, susie performs best. And survival svb has very high FDR.

2. When X are close to independent, susie is slightly worse than the other two methods. 


```{r}
calculate_tpr_vs_fdr <- function(pip, is_effect, ts){
  res <- matrix(NA, nrow = length(ts), ncol = 2)
  colnames(res) = c("tpr", "fdr")
  for (i in 1:length(ts)){
    pred_pos = pip >= ts[i]
    tp = pip >= ts[i] & is_effect == 1
    fp = pip >= ts[i] & is_effect == 0
    tpr = sum(tp)/sum(is_effect)
    fdr = sum(fp)/sum(pred_pos)
    res[i, ] = c(tpr, fdr)
  }
  return(res)
}

```

```{r}
susie = readRDS("./data/susie.rds")
survsvb = readRDS("./data/survsvb.rds")
bvsnlp = readRDS("./data/bvsnlp.rds")
```

```{r}
p = 1e3
indx = which(susie$simulate.cor_type == "real")
pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))
pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))

is_effect = c()
for (i in indx){
    init = rep(0, p)
    effect_indx = susie$simulate.effect_indx[[i]]
    init[effect_indx] = 1
    is_effect = c(is_effect, init)
}

```

```{r}
ts = seq(from = 0, to = 1, by = 0.01)
res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
```


```{r}
plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0,1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
     main = "Effect number 1-3 & X highly correlated")
lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)

points(res.susie[which(ts == 0.9),2], res.susie[which(ts == 0.9), 1])
points(res.svb[which(ts == 0.9),2], res.svb[which(ts == 0.9), 1])
points(res.bvsnlp[which(ts == 0.9),2], res.bvsnlp[which(ts == 0.9), 1])

legend("topleft", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
```
The dots indicate PIP threshold = 0.9

```{r}
p = 1e3
indx = which(susie$simulate.cor_type == "independent")
pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))
pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))

is_effect = c()
for (i in indx){
    init = rep(0, p)
    effect_indx = susie$simulate.effect_indx[[i]]
    init[effect_indx] = 1
    is_effect = c(is_effect, init)
}

```

```{r}
ts = seq(from = 0, to = 1, by = 0.01)
res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
```


```{r}
plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 0.2), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
     main = "Effect number 1-3 & X near independent")
lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)

points(res.susie[which(ts == 0.9),2], res.susie[which(ts == 0.9), 1])
points(res.svb[which(ts == 0.9),2], res.svb[which(ts == 0.9), 1])
points(res.bvsnlp[which(ts == 0.9),2], res.bvsnlp[which(ts == 0.9), 1])

legend("topleft", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
```
The dots indicate PIP threshold = 0.9
