---
title: "Compare PIPs"
author: "Yunqi Yang"
date: "4/6/2023"
output: html_document
---

### Description:

Simulation results based on real genotype data from GTEx. I took the SNPs for a certain gene, Thyroid.ENSG00000132855. Therefore, sample size $n=584$, and I choose variable number $p=1000$.

#### Steps for simulating survival data:

1. Select a random point on the genome, indx_start. Then the predictors are from [indx_start:(indx_start + p -1)]. 

2. Select effect variables. The number of effects are 1,2,3. Select SNPs that have correlation between (0.9 - 0.99) with other SNPs. 

3. Two correlation types: real correlation and independent. In independent, I simply permute each variable values. The max correlation can be 0.3, but a lot of correlation will be close to 0. 

4. Effect size $b\sim N(0,1)$.

5. Simulate survival time using exponential survival model. 


#### Fitting susie:

1. Set L = true number of effects. 

2. Set prior variance to the true prior variance. 


#### Method comparison:

1. survival.svb:https://doi.org/10.1093/bioinformatics/btac416

2. bvsnlp: https://yunqiyang0215.github.io/survival-susie/bvsnlp.html


#### **Conclusion**:

1. When using the real correlation matrix, susie result is very different from survival.svb. And survival.svb seems to have better results, as capturing more effects. 

2. In other scenarios, results are more consistent across methods. 

```{r knitr, echo=FALSE}
knitr::opts_chunk$set(comment = "#",results = "hold",collapse = TRUE,
                      fig.align = "center", warning = FALSE)
```

```{r}
compare_pip <- function(pip1, pip2, effect_indx, dat_indx, p){
  is_effect = c()
  for (i in dat_indx){
    init = rep(0, p)
    indx = effect_indx[[i]]
    init[indx] = 1
    is_effect = c(is_effect, init)
  }
  res = data.frame(cbind(pip1, pip2, is_effect))
}
```

```{r}
plot_pip <- function(res.pip, labs, main){
  pch = ifelse(res.pip$is_effect == 1, 16, 1)
  cex = ifelse(res.pip$is_effect == 1, 1, 0.5)
  col = ifelse(res.pip$is_effect == 1, "red", "dark grey")
  plot(res.pip$pip1, res.pip$pip2, col = col, pch = pch, cex = cex, 
       xlab = labs[1], ylab = labs[2], main = main)
}
```

```{r}
susie = readRDS("./data/susie.rds")
survsvb = readRDS("./data/survsvb.rds")
bvsnlp = readRDS("./data/bvsnlp.rds")
```

### 1. Check if each row stores the same simulated data
```{r}
sum(bvsnlp$simulate.cor_type != susie$simulate.cor_type)
sum(bvsnlp$simulate.num_effect != susie$simulate.num_effect)
sum(unlist(bvsnlp$simulate.effect_indx) != unlist(susie$simulate.effect_indx))
sum(unlist(bvsnlp$simulate.effect_size) != unlist(susie$simulate.effect_size))
```

### 1. Use the real correlation of SNPs
```{r fig.width= 10, fig.height=5}
par(mfrow = c(1,3))
for (num in 1:3){
  indx = which(susie$simulate.cor_type == "real" & susie$simulate.num_effect == num)
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))

  res.pip = compare_pip(pip.susie, pip.survsvb, susie$simulate.effect_indx, dat_indx = indx, p =1e3)
  plot_pip(res.pip, labs = c("susie", "survival.svb"), main = paste0(num, " effect, max corr=0.9-0.99"))
}
```

```{r fig.width= 10, fig.height=5}
par(mfrow = c(1,3))
for (num in 1:3){
  indx = which(susie$simulate.cor_type == "real" & susie$simulate.num_effect == num)
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))

  res.pip = compare_pip(pip.susie, pip.bvsnlp, susie$simulate.effect_indx, dat_indx = indx, p =1e3)
  plot_pip(res.pip, labs = c("susie", "bvsnlp"), main = paste0(num, " effect, max corr=0.9-0.99"))
  abline(a = 0, b = 1, lty = 2, col = "blue")
}
```

### 2. Near independent SNPs
```{r fig.width= 10, fig.height=5}
par(mfrow = c(1,3))
for (num in 1:3){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.num_effect == num)
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))

  res.pip = compare_pip(pip.susie, pip.survsvb, susie$simulate.effect_indx, dat_indx = indx, p =1e3)
  plot_pip(res.pip, labs = c("susie", "survival.svb"), main = paste0(num, " effect, X close to independent"))
  abline(a = 0, b = 1, lty = 2, col = "blue")
}
```

```{r fig.width= 10, fig.height=5}
par(mfrow = c(1,3))
for (num in 1:3){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.num_effect == num)
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))

  res.pip = compare_pip(pip.susie, pip.bvsnlp, susie$simulate.effect_indx, dat_indx = indx, p =1e3)
  plot_pip(res.pip, labs = c("susie", "bvsnlp"), main = paste0(num, " effect, X close to independent"))
  abline(a = 0, b = 1, lty = 2, col = "blue")
}

```

### Compare susie pip and bvsnlp pip:

```{r}
indx = which(susie$simulate.cor_type == "independent" & susie$simulate.num_effect == 1)
pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
res.pip = compare_pip(pip.susie, pip.bvsnlp, susie$simulate.effect_indx, dat_indx = indx, p =1e3)
```

```{r}
res.pip[res.pip$pip1 > 0.8 & res.pip$is_effect == 0, ]
```

Susie with false positive, high PIPs and bvsnlp with low PIPs: 
```{r}
example_indx = indx[64]
example_indx
```

```{r}
pip.susie = unlist(lapply(example_indx, function(x) susie$susie.pip[[x]]))
pip.bvsnlp = unlist(lapply(example_indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
res.pip = compare_pip(pip.susie, pip.bvsnlp, susie$simulate.effect_indx, dat_indx = example_indx, p =1e3)
res.pip[res.pip$pip1 > 0.8 & res.pip$is_effect == 0, ]
```


