---
title: "numerical_int_gauss_quad"
author: "Yunqi Yang"
date: "2/13/2024"
output: html_document
---

We know we can use partial likelihood to compute BF from here: https://yunqiyang0215.github.io/survival-susie/numerical_integration.html. Here we use Gaussian Quadrature approach (implemented by Karl) to do numerical integration to compute BF.

$$
  \begin{split}
BF&=\frac{P(D|H_1)}{P(D|H_0)}=\frac{\int Lp(b)p(b)db}{Lp(b=0)}\\
Lp(b)&=\prod_{i=1}^n\{\frac{\exp(x_ib)}{\sum_{j\in R(y_i)}\exp(x_jb)} \}^{\delta_i}
\end{split}
$$


Let $f(b)=L_p(b)p(b)$, which is the input $f(x)$ in the code. *mu* and *sd* are the posterior mean and standard error of $b$ if we approximate $L_p(b)\approx N(b;\hat b, s^2)$, where $\hat b$ is the maximize partial likelihood estimate and $s^2$ is the corresponding standard error. 


```{r}
#' Gaussian quadrature
#'
#' approximate log(\int f(x)dx) by quadrature on \int f(x)/p(x) p(x)
#' where p(x) is a normal density, we approximate this integral
#' with a gaussian quadrature rule via `statmod::gauss.quad.prob`
#' done on a log scale for numerical stability (f is positive)
#' the goal is to choose p(x) such that f(x)/p(x) is not too variable near the mode
#' @param mu mean of gaussian quadrature
#' @param sd sd of gaussian quadrature
gauss_quad2 <- function(log_f, mu, sd, n=32){
  q <- statmod::gauss.quad.prob(n, dist='normal', mu=mu, sigma=sd)
  log_integrand <- function(x){log_f(x) - dnorm(x, mu, sd, log=T)}
  return(matrixStats::logSumExp(log_integrand(q$nodes) + log(q$weights)))
}
```

```{r}
# @param b: the effect size
# @param surT: a Surv() object, containing time and status
# @param x: covariate
get_partial_loglik <- function(b, survT, x){
  ll.partial = logLik(coxph(survT ~ x, init = c(b), control=list('iter.max'= 0, timefix=FALSE)))
  return(as.numeric(ll.partial))
}

# @param b: the effect size
# @param surT: a Surv() object, containing time and status
# @param x: covariate
# @param sig0: prior standard deviation
get_log_f <- function(b, survT, x, sig0){
  partial_ll = get_partial_loglik(b, survT, x)
  return(partial_ll + dnorm(b, 0, sd = sig0, log = TRUE))
}

```

```{r}
library(survival)
library(statmod)
source("./code/surv_susie_helper.R")
source("./code/sim_surv.R")
```


### Simulate data

```{r}
set.seed(1)
# simulate 2 variables 
n = 100
X = cbind(rbinom(n, size = 2, prob = 0.3), rbinom(n, size = 1, prob = 0.3))
# the first element of b is for intercept
b = c(1, 1, 0)
censor_lvl = 0.7
dat <- sim_surv_exp(b, X, censor_lvl)
survT <- Surv(dat$y[,1], dat$y[,2])

```

```{r}
prior_sd = 1
res = surv_uni_fun(x = dat$X[,1], y = survT, o = rep(0, n), prior_variance = prior_sd^2)
mu = res$mu
sd = sqrt(res$var)
```

```{r}
log_f <- function(b) get_log_f(b, survT, x = dat$X[,1], sig0 = prior_sd)
lbf.numerical = gauss_quad2(Vectorize(log_f), mu, sd, n=32) - get_partial_loglik(b = 0, survT, dat$X[,1])
```

Result seems very reasonable!

```{r}
print(lbf.numerical)
print(res$lbf)
```

