---
title: "Check the distribution of zscores in coxph"
author: "Yunqi Yang"
date: "2/9/2023"
output: html_document
---

## Description:

Here I first simulate survival data from exponential distribution **without censoring**. I simulate one variable and the true model is the null model, that is, the variable is not associated with survival time $T$. From fitting the cox proportional hazard model on the data, we characterize the distribution of z-scores. 


### More background on survival analysis:

**1. To characterize the distribution of survival time $T$**, people often consider 4 functions, which are interrelated. Their relationships can be derived, here I just summarize the facts:

(1) probability density function, f(t). $f(t)=\lambda(t)S(t)$

(2) survival function, S(t). $S(t)=1-F(t)$

(3) hazard function, $\lambda(t)$. $\lambda(t)=f(t)/S(t)=-\frac{\partial}{\partial t}\log S(t)$.

(4) cumulative hazard function, $\Lambda(t)$. $\Lambda(t)=\int_0^t\lambda(s)d s$.


**2. Parametric distribution for $T$**: exponential distribution

In exponential distribution, the hazard (instantaneous rate at time $t$) is constant over time. 
$$
\begin{split}
f(t)&=\frac{1}{\mu}\exp\{-t/\mu\}\\
\lambda(t)&=\frac{1}{\mu}
\end{split}
$$

**3. Model the mean of survival time**:

Remember in exponential distribution, $E(T)=\mu$. So we can model $\log\mu$ by linear combinations of variables. $i$ indexes an individual. Let's call this *generative model*, because I simulate $T_i$ from $\exp(\mu_i)$.

$$
\begin{split}
\log T_i &= \log E(T_i) + \epsilon_i=\log\mu_i+\epsilon_i\\
\log\mu_i &=\beta_0+\beta ^Tx_i
\end{split}
$$

**4. Cox proportional hazard model**:

Cox proportional hazard (cox-ph) model is a model for the hazard $\lambda(t)$, not the survival time. It has a form of the following, where $\lambda_0$ is the baseline hazard and $x$ (length-p vector) denotes attributes of an individual that increase/decrease his hazard. For example, high BMI may increase the hazard of getting diabetes. 

$$
\lambda_i(t)=\lambda_0\exp\{\alpha_0+\alpha^Tx_i\}
$$

Cox-ph model is a semi-parametric model, where the baseline hazard function $\lambda_0$ can be left unspecified. The inference carries out only in $\alpha_0, \alpha$. 


**5. Relationship between our generative model and cox-ph**:

As mentioned above, when assuming $T$ follows exponential distribution, we have $\lambda=1/\mu$, again this means the hazard is constant over time. Here I mainly want to show the relationship between $\alpha$ in coxph and $\beta$ in our generative model.

$$
\begin{split}
\lambda_i=1/\mu_i&=\frac{1}{\exp\{\beta_0+\beta^Tx_i\}}=\frac{1}{\mu_0\exp\{\beta^Tx_i\}}\\
&=\lambda_0\exp\{-\beta^Tx_i\}=\lambda_0\exp\{-\alpha^Tx_i\}
\end{split}
$$

Therefore, $\alpha = -\beta$. That is to say, if I generate a model with $\beta_1=1$ and fit the survival data to cox-ph model, the coefficient I get is $\alpha$, which is -1. 



```{r}
library(mvtnorm)
library(survival)
# Modified Karl's code for intercept part
devtools::load_all("/Users/nicholeyang/Desktop/logisticsusie")
```

### Step 1: simulate data from the null model. 

Let's do the simplest case, where we only have one variable and is not associated with survival time. 

```{r}
# Here we use parametric model to simulate data with survival time,
# assuming survival time is exponentially distributed. 
# We first simulate the mean of exponential from linear combinations
# of variables, and then simulate survival time. 
# T\sim 1/u*exp(-t/u), and the true model is:
# log(T) = \mu + e = b0 + Xb + e
# @param b: vector of length (p+1) for true effect size, include intercept.
# @param X: variable matrix of size n by p. 
# @param status: censoring status. 1 = censored, 2 = event observed. 
sim_dat <- function(b, X){
  n = nrow(X)
  p = ncol(X)
  mu <- exp(cbind(rep(1,n), X) %*% b)
  surT <- rexp(n, rate = 1/mu)
  dat <- data.frame(cbind(surT, X))
  x.name <- unlist(lapply(1:p, function(i) paste0("x", i)))
  names(dat) = c("surT", x.name)
  dat$status <- rep(2, n)
  return(dat)
}

```

```{r}
set.seed(1)
n = 200
b = c(1, 0) # True beta coefficients, first element is intercept. 
X = as.matrix(rnorm(n, mean = 3, sd = 2))
dat = sim_dat(b, X)
hist(dat$surT, breaks = 20)
```

### Step 2: fit cox-ph model
```{r}
## Create  survival object. status == 2 is death
dat$y <- with(dat, Surv(surT, status == 2))
cox1 <- coxph(y ~ x1, data =  dat)
summary(cox1)
```
### Step 3: simulate multiple data replicates and fit coxph, plot the distribution of z-scores
```{r}
seeds = c(1:5e3)
n = 200
b = c(1, 0)
zscores <- rep(NA, 5e3)

for (seed in seeds){
  
  set.seed(seed)
  X = as.matrix(rnorm(n, mean = 3, sd = 2))
  dat = sim_dat(b, X)
  
  ## Create  survival object. status == 2 is death
  dat$y <- with(dat, Surv(surT, status == 2))
  cox <- coxph(y ~ x1, data =  dat)
  zscores[seed] = coef(summary(cox))[,"z"]
  
}

```

```{r}
x = seq(-3, 3, length.out = 1e3)
hist(zscores, probability = TRUE)
# The red line is the density of N(0,1)
curve(dnorm(x, mean = 0, sd = 1), add = TRUE, col = "red")

```

