---
title: "compute_bf"
author: "Yunqi Yang"
date: "6/27/2023"
output: html_document
---


### Description: using quadrature method to compute Bayes factor in survival susie model. 

The standard BF for comparing the hypothesis of having one predictor vs. the null model:
$$

\frac{P(D|H_1)}{P(D|H_0)}=\frac{\int\int p(y,\delta|x,b, h_0)p(b)dbdh_0}{\int p(y,\delta|x,b=0,h_0)dh_0}

$$

Based on the paper, "A Bayesian Justification of Cox’s Partial Likelihood" by Sinha, D. and others, we can safely use the partial likelihood to compute Bayes factor. That is,

$$
\begin{split}
\frac{P(D|H_1)}{P(D|H_0)}&=\frac{\int Lp(b)p(b)db}{Lp(b=0)}\\
Lp(b)&=\prod_{i=1}^n\{\frac{\exp(x_ib)}{\sum_{j\in R(y_i)}\exp(x_jb)} \}^{\delta_i}
\end{split}
$$

```{r}
# Function to simulate survival time under exponential model. That is,
# assuming survival time is exponentially distributed.
# lambda(t) = lambda*exp(b0 + Xb). S(t) = exp(-lambda*t),
# F(t) = 1- S(t) \sim Unif(0,1). Therefore, t = log(1-runif(0,1))/-exp(b0+Xb).
# For censored objects, we simulate the censoring time by rescale the actual survival time.
# @param b: vector of length (p+1) for true effect size, include intercept.
# @param X: variable matrix of size n by p.
# @param censor_lvl: a constant from [0,1], indicating the censoring level in the data.
# @return  dat: a dataframe that contains `y`, `x` and `status`.
# `status`: censoring status: 0 = censored, 1 = event observed. See Surv() in library(survival)
sim_surv_exp <- function(b, X, censor_lvl){
  n = nrow(X)
  p = ncol(X)
  dat = list()

  status <- ifelse(runif(n) > censor_lvl, 1, 0)
  lambda <- exp(cbind(rep(1,n), X) %*% b)
  surT <- log(1 - runif(n)) /(-lambda)
  # rescale censored subject to get observed time
  surT[status == 0] = surT[status == 0] * runif(sum(status == 0))

  y = cbind(surT, status)
  colnames(y) = c("time", "status")
  colnames(X) <- unlist(lapply(1:p, function(i) paste0("x", i)))
  dat[["X"]] = X
  dat[["y"]] = y
  return(dat)
}

```

```{r}
library(survival)
library(cubature)
source("./code/surv_susie_helper.R")
```


### Simulate data
```{r}
set.seed(1)
# simulate 2 variables 
n = 100
X = cbind(rbinom(n, size = 2, prob = 0.3), rbinom(n, size = 1, prob = 0.3))
# the first element of b is for intercept
b = c(1, 1, 0)
censor_lvl = 0.7
dat <- sim_surv_exp(b, X, censor_lvl)
survT <- Surv(dat$y[,1], dat$y[,2])
```


### Use numerical integration 
```{r code-numerical-integration}
# @param b: the effect size
# @param surT: a Surv() object, containing time and status
# @param x: covariate
get_partial_lik <- function(b, survT, x){
  partial.loglik = logLik(coxph(survT ~ x, init = c(b), control=list('iter.max'= 0, timefix=FALSE)))
  partial.lik = exp(partial.loglik) # handling ties...
  return(as.numeric(partial.lik))
}

integrand <- function(b, survT, x, prior_sd){
  val = get_partial_lik(b, survT, x)*dnorm(b, mean = 0, sd = prior_sd)
  return(val)
}
```

```{r numerical-integration-res}
prior_sd = 1
res = cubintegrate(f = integrand, survT = survT, x = dat$X[,1], prior_sd = 1, lower = -Inf, upper = Inf, method = "hcubature")
res
prob.h0 = get_partial_lik(b = 0, survT, dat$X[,1])
lbf.numerical = log(res$integral) - log(prob.h0)
```

```{r}
abf <-surv_uni_fun(x = dat$X[,1], y = survT, o = rep(0, n), prior_variance = prior_sd^2)$lbf
```

```{r}
lbf.numerical
abf
```

