---
title: "sim_survival"
author: "Yunqi Yang"
date: "2/1/2023"
output: html_document
---

## Description:
Simulate time-to-event data based on exponential model. And fit proportional hazard model to data. Let's first simulate data without **censoring**. The full explanation about simulation is at: https://yunqiyang0215.github.io/survival-susie/sim_survival.html


Here I try different correlations between $x_1$ and $x_2$ where $x_1$ has an effect on survival time, and $x_2$ doesn't have. That is, the true model is: $\log T_i = \beta_0+\beta_1x_{i1} +\epsilon_i$. 

Results: 
(1) For r = 0.9, and r = 0.8. Susie seems work marginally. 
(2) For r = 0.7, Susie gives rediculous results. Seems something wrong in the code. 
```{r}
library(mvtnorm)
library(survival)
# Modified Karl's code for intercept part
devtools::load_all("/Users/nicholeyang/Desktop/logisticsusie")
```

### Simulation functions:
```{r}
# Here we use parametric model to simulate data with survival time,
# assuming survival time is exponentially distributed. 
# We first simulate the mean of exponential from linear combinations
# of variables, and then simulate survival time. 
# T\sim 1/u*exp(-t/u), and the true model is:
# log(T) = \mu + e = b0 + Xb + e
# @param b: vector of length (p+1) for true effect size, include intercept.
# @param X: variable matrix of size n by p. 
# @param status: censoring status. 1 = censored, 2 = event observed. 
sim_dat <- function(b, X){
  n = nrow(X)
  mu <- exp(cbind(rep(1,n), X) %*% b)
  surT <- rexp(n, rate = 1/mu)
  dat <- data.frame(cbind(surT, X))
  names(dat) = c("surT", "x1", "x2")
  dat$status <- rep(2, n)
  return(dat)
}
```

### Functions for running IBSS
```{r}
# Function to calculate approximate BF based on Wakefield approximation
# @param z: zscore of the regression coefficient
# @param s: standard deviation of the estimated coefficient
compute_abf <- function(z, s, prior_variance){
  abf <- sqrt(s^2/(s^2+prior_variance))*exp(z^2/2*(prior_variance/(s^2+prior_variance)))
  return(abf)
}


compute_approx_post_var <- function(z, s, prior_variance){
  post_var <- 1/(1/s^2 + 1/prior_variance)
  return(post_var)
}

# @param post_var: posterior variance
# @param s: standard deviation of the estimated coefficient
# @param bhat: estimated beta effect
compute_approx_post_mean <- function(post_var, s, bhat){
  mu <- post_var/(s^2)*bhat
  return(mu)
}
```

```{r}
surv_uni_fun <- function(x, y, o, prior_variance, estimate_intercept = 0, ...){
  fit <- coxph(y~ x + o)
  bhat <- summary(fit)$coefficients[1, 1]
  sd <- summary(fit)$coefficients[1, 3]
  zscore <- summary(fit)$coefficients[1, 4]
  
  bf <- compute_abf(zscore, sd, prior_variance)
  var <- compute_approx_post_var(zscore, sd, prior_variance)
  mu <- compute_approx_post_mean(var, sd, bhat)
  lbf <- log(bf)
  return(list(mu = mu, var=var, lbf=lbf, intercept=0))
}

fit_coxph <- ser_from_univariate(surv_uni_fun)

```

#### Sim1: r = 0.9. 
```{r}
set.seed(2)
r = 0.9
n <- 50
b <- c(1, 3, 0)
X <- rmvnorm(n, sigma = matrix(c(1, r, r, 1), ncol = 2, nrow = 2))
dat <- sim_dat(b, X)
hist(dat$surT, breaks = 20)
```
```{r}
# Fit cox ph
## Create  survival object. status == 2 is death
dat$y <- with(dat, Surv(surT, status == 2))
cox <- coxph(y ~ x1 + x2, data =  dat)
summary(cox)
```

```{r}
X = as.matrix(dat[, c(2:3)])
y = dat$y

fit <- ibss_from_ser(X, y, L = 10, prior_variance = 1., prior_weights = rep(1/2, 2), tol = 1e-3, maxit = 100, estimate_intercept = TRUE, ser_function = fit_coxph)
fit$alpha
```


#### Sim2: r = 0.8 
```{r}
set.seed(2)
r = 0.8
n <- 50
b <- c(1, 3, 0)
X <- rmvnorm(n, sigma = matrix(c(1, r, r, 1), ncol = 2, nrow = 2))
dat <- sim_dat(b, X)
hist(dat$surT, breaks = 20)
```

```{r}
# Fit cox ph
## Create  survival object. status == 2 is death
dat$y <- with(dat, Surv(surT, status == 2))
cox <- coxph(y ~ x1 + x2, data =  dat)
summary(cox)
```

```{r}
X = as.matrix(dat[, c(2:3)])
y = dat$y

fit <- ibss_from_ser(X, y, L = 10, prior_variance = 1., prior_weights = rep(1/2, 2), tol = 1e-3, maxit = 100, estimate_intercept = TRUE, ser_function = fit_coxph)
fit$alpha
```





#### Sim3: r = 0.7
```{r}
set.seed(2)
r = 0.7
n <- 50
b <- c(1, 3, 0)
X <- rmvnorm(n, sigma = matrix(c(1, r, r, 1), ncol = 2, nrow = 2))
dat <- sim_dat(b, X)
hist(dat$surT, breaks = 20)
```

```{r}
# Fit cox ph
## Create  survival object. status == 2 is death
dat$y <- with(dat, Surv(surT, status == 2))
cox <- coxph(y ~ x1 + x2, data =  dat)
summary(cox)
```

```{r}
X = as.matrix(dat[, c(2:3)])
y = dat$y

fit <- ibss_from_ser(X, y, L = 10, prior_variance = 1., prior_weights = rep(1/2, 2), tol = 1e-3, maxit = 100, estimate_intercept = TRUE, ser_function = fit_coxph)
fit$alpha
```

