---
title: "Fit coxph susie on pdac gene expression"
output: html_document
date: '2024-01-31'
---

1. Model fitting script: /project2/mstephens/yunqiyang/surv-susie/realdata/pdac_dat/fit_pdac.R

2. Result: coxph-SuSIE doesn't find out any signal, while bvsnlp finds 2 genes. However, the 2 genes BVSNLP finds doesn't have significant p-value or large Bayes factor after controlling for stage variable. 

```{r}
library(survival)
library(susieR)
devtools::load_all("/project2/mstephens/yunqiyang/surv-susie/logisticsusie")
source("/project2/mstephens/yunqiyang/surv-susie/survival-susie/code/surv_susie_helper.R")
```

```{r}
res = readRDS("/project2/mstephens/yunqiyang/surv-susie/realdata/pdac_dat/fit_pdac.rds")
dat = readRDS("/project2/mstephens/yunqiyang/surv-susie/realdata/pdac_dat/dat_pdac.rds")
expression = dat[[2]]
fit.susie = res[[1]]
fit.bvsnlp = res[[2]]
```

### 1. BVSNLP result:

```{r}
p = ncol(expression)
effect_estimate = rep(0, p)
effect_estimate[fit.bvsnlp$HPM] = fit.bvsnlp$beta_hat
pip = fit.bvsnlp$inc_probs
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:10]
colnames(expression)[effect_estimate != 0]
```


### 2. Susie result:

Susie finds no credible sets even change min_abs_corr = 0.1. 

```{r}

pip <- logisticsusie:::get_pip(fit.susie$alpha)
pip.sorted = sort(pip, decreasing = TRUE)
pip.sorted[1:5]
class(fit.susie) = "susie"
cs <- susie_get_cs(fit.susie, as.matrix(expression), min_abs_corr = 0.1)
print(cs)
iter = fit.susie$iter
print(iter)
```

### 3. Fit coxph and compute BF on the two genes BVSNLP found:

```{r}
y = Surv(dat[[1]]$futime, dat[[1]]$event)
fit1 = coxph(y ~ dat[[1]]$stage + dat[[2]][, "GTPBP6"])
fit2 = coxph(y ~ dat[[1]]$stage + dat[[2]][, "ACSF3"])
summary(fit1)
summary(fit2)
```

```{r}
# adjust for stage variable
cov_fit <- coxph(y ~ dat[[1]]$stage)
fixed <- predict(cov_fit)
      
lbf1 = surv_uni_fun(x = dat[[2]][, "GTPBP6"], y = y, o = fixed, prior_variance = 0.1, estimate_intercept = 0)$lbf
lbf2 = surv_uni_fun(x = dat[[2]][, "ACSF3"], y = y, o = fixed, prior_variance = 0.1, estimate_intercept = 0)$lbf
print(lbf1)
print(lbf2)
```




