---
title: "Power vs. fdr"
author: "Yunqi Yang"
date: "4/17/2023"
output: html_document
---

### Description:

New simulation results, comparing power vs. FDR across 3 methods. I vary the threshold for claiming effect variables based on marginal PIP value. 

**Difference between old and new simulation**:

1. New simulation removed SNPs with MAF < 5%. 

2. Added the null data, no effect variable. 

3. Added censoring to data simulation. 

4. Force the effect variable not to be the first varaible.

5. Added the procedure for estimating prior variance. 

Other simulation settings, see here: https://yunqiyang0215.github.io/survival-susie/compare_pip.html


**Conclusion**:

1. When X are highly correlated, susie performs best. 

2. When X are independent, and without null model data, all methods are similar. 

3. survival.svb and bvsnlp doesn't perform well after adding data from null model.

4. When censoring level = 0.4, susie performance starts to get worse. 

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center")
```

```{r}
calculate_tpr_vs_fdr <- function(pip, is_effect, ts){
  res <- matrix(NA, nrow = length(ts), ncol = 2)
  colnames(res) = c("tpr", "fdr")
  for (i in 1:length(ts)){
    pred_pos = pip >= ts[i]
    tp = pip >= ts[i] & is_effect == 1
    fp = pip >= ts[i] & is_effect == 0
    tpr = sum(tp)/sum(is_effect)
    fdr = sum(fp)/sum(pred_pos)
    res[i, ] = c(tpr, fdr)
  }
  return(res)
}

```

```{r}
susie = readRDS("./data/dsc2/susie.rds")
survsvb = readRDS("./data/dsc2/survsvb.rds")
bvsnlp = readRDS("./data/dsc2/bvsnlp.rds")
```

```{r}
names(susie)
```

### 1. Results using real correlation structure from data

```{r, fig.width=12, fig.height=5}
par(mfrow = c(1,3))
censor_lvl = c(0, 0.2, 0.4)
for (i in 1:3){
  indx = which(susie$simulate.cor_type == "real" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 1, by = 0.01)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0,1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Effect 0-3 & max cor(X)=0.9-0.99", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  
  points(res.susie[96,2], res.susie[96, 1])
  points(res.svb[96,2], res.svb[96, 1])
  points(res.bvsnlp[96,2], res.bvsnlp[96, 1])
  
  legend("topleft", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```
The dots indicate PIP threshold = 0.95


### 2. Results using independent X, without data from null model. 

```{r, fig.width=12, fig.height=5}
par(mfrow = c(1,3))
censor_lvl = c(0, 0.2, 0.4)

for (i in 1:3){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i] & susie$simulate.num_effect != 0)
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  
  ts = seq(from = 0, to = 1, by = 0.01)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 0.1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Effect 1-3 & X independent", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  points(res.susie[96,2], res.susie[96, 1])
  points(res.svb[96,2], res.svb[96, 1])
  points(res.bvsnlp[96,2], res.bvsnlp[96, 1])
  legend("bottomright", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```

The dots indicate PIP threshold = 0.95.

```{r}
sum(is_effect == 1)
which(pip.susie > 0.95 & is_effect == 0)
sum(pip.susie > 0.95 & is_effect == 1)
sum(pip.susie > 0.99 & is_effect == 1)
sum(pip.susie ==1 & is_effect == 1)

```

```{r}
which(pip.bvsnlp > 0.95 & is_effect == 0)
sum(pip.bvsnlp > 0.95 & is_effect == 1)
sum(pip.bvsnlp > 0.99 & is_effect == 1)
sum(pip.bvsnlp ==1 & is_effect == 1)

```



### 3. Results using independent X, with data from null model. 

```{r, fig.width=12, fig.height=5}
par(mfrow = c(1,3))
censor_lvl = c(0, 0.2, 0.4)

for (i in 1:3){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  
  ts = seq(from = 0, to = 1, by = 0.01)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Effect 0-3 & X independent", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  points(res.susie[96,2], res.susie[96, 1])
  points(res.svb[96,2], res.svb[96, 1])
  points(res.bvsnlp[96,2], res.bvsnlp[96, 1])
  legend("bottomright", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```

The dots indicate PIP threshold = 0.95.


```{r, fig.width=12, fig.height=5}
par(mfrow = c(1,3))
censor_lvl = c(0, 0.2, 0.4)

for (i in 1:3){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.survsvb = unlist(lapply(indx, function(x) survsvb$survivalsvb.pip[[x]]))
  pip.bvsnlp = unlist(lapply(indx, function(x) bvsnlp$bvsnlp.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  
  ts = seq(from = 0, to = 1, by = 0.01)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.svb = calculate_tpr_vs_fdr(pip.survsvb, is_effect, ts)
  res.bvsnlp = calculate_tpr_vs_fdr(pip.bvsnlp, is_effect, ts)
  
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0, 0.1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Effect 0-3 & X independent", ",censor=", censor_lvl[i]))
  lines(res.svb[,2], res.svb[,1], type = "l", col = 2)
  lines(res.bvsnlp[,2], res.bvsnlp[,1], type = "l", col = 3)
  points(res.susie[96,2], res.susie[96, 1])
  points(res.svb[96,2], res.svb[96, 1])
  points(res.bvsnlp[96,2], res.bvsnlp[96, 1])
  legend("bottomright", legend = c("susie", "survival.svb", "bvsnlp"), col = c(1,2,3), lty = 1)
}

```



### 4. PIP histogram of survival.svb

**4.1** Under real correlation
The null data contains 300 replicates, the non-null data has 900 replicates. 
On average, the non-null data contains 2 effects per data.

```{r}
indx1 = which(survsvb$simulate.cor_type == "real" & survsvb$simulate.num_effect == 0)
pip1 = unlist(lapply(indx1, function(x) survsvb$survivalsvb.pip[[x]]))
sum(pip1 > 0.95)
length(indx1)

indx2 = which(survsvb$simulate.cor_type == "real" & survsvb$simulate.num_effect != 0)
pip2 = unlist(lapply(indx2, function(x) survsvb$survivalsvb.pip[[x]]))
sum(pip2 > 0.95)
length(indx2)
```


```{r}
par(mfrow = c(1,2))
hist(pip1, xlab = "pip", main = "Null data: X high correlation")
hist(pip2, xlab = "pip", main = "Non-null data:  X high correlation")
```

**4.2** Under independent correlation
The null data contains 300 replicates, the non-null data has 900 replicates. 
On average, the non-null data contains 2 effects per data.

```{r}
indx1 = which(survsvb$simulate.cor_type == "independent" & survsvb$simulate.num_effect == 0)
pip1 = unlist(lapply(indx1, function(x) survsvb$survivalsvb.pip[[x]]))
sum(pip1 > 0.95)
length(indx1)

indx2 = which(survsvb$simulate.cor_type == "independent" & survsvb$simulate.num_effect != 0)
pip2 = unlist(lapply(indx2, function(x) survsvb$survivalsvb.pip[[x]]))
sum(pip2 > 0.95)
length(indx2)
```

```{r}
par(mfrow = c(1,2))
hist(pip1, xlab = "pip", main = "Null data: X independent")
hist(pip2, xlab = "pip", main = "Non-null data: X independent")
```


### 5. PIP histogram of susie

**4.1** Under real correlation
The null data contains 300 replicates, the non-null data has 900 replicates. 
On average, the non-null data contains 2 effects per data.

```{r}
indx1 = which(susie$simulate.cor_type == "real" & susie$simulate.num_effect == 0)
pip1 = unlist(lapply(indx1, function(x) susie$susie.pip[[x]]))
sum(pip1 > 0.95)
length(indx1)

indx2 = which(susie$simulate.cor_type == "real" & susie$simulate.num_effect != 0)
pip2 = unlist(lapply(indx2, function(x) susie$susie.pip[[x]]))
sum(pip2 > 0.95)
length(indx2)
```


```{r}
par(mfrow = c(1,2))
hist(pip1, probability = TRUE,  xlab = "pip", main = "Null data: X high correlation")
hist(pip2, probability = TRUE, xlab = "pip", main = "Non-null data:  X high correlation")
```

**4.2** Under independent correlation
The null data contains 300 replicates, the non-null data has 900 replicates. 
On average, the non-null data contains 2 effects per data.

```{r}
indx1 = which(susie$simulate.cor_type == "independent" & susie$simulate.num_effect == 0)
pip1 = unlist(lapply(indx1, function(x) susie$susie.pip[[x]]))
sum(pip1 > 0.95)
length(indx1)

indx2 = which(susie$simulate.cor_type == "independent" & susie$simulate.num_effect != 0)
pip2 = unlist(lapply(indx2, function(x) susie$susie.pip[[x]]))
sum(pip2 > 0.95)
length(indx2)
```


```{r}
par(mfrow = c(1,2))
hist(pip1, probability = TRUE,  xlab = "pip", main = "Null data: X high correlation")
hist(pip2, probability = TRUE, xlab = "pip", main = "Non-null data:  X high correlation")
```
