---
title: "Apply SER module"
author: "Yunqi Yang"
date: "2/5/2023"
output: html_document
---
### Description:

I simulated data in 5 different scenarios, and apply both coxph model and susie procedure. The data simulation procedure is available at https://yunqiyang0215.github.io/survival-susie/sim_survival.html


#### The Wakefield approximation:
$$
ABF(H_1/H_0)=\sqrt\frac{V}{V+W}\exp\{\frac{z^2}{2}\frac{W}{V+W}\},
$$
where $V$ is the variance of estimated regression coefficient, and $W$ is variance in the normal prior, $N(0,W)$. 

```{r}
# Function to calculate approximate BF based on Wakefield approximation
# @param z: zscore of the regression coefficient
# @param s: standard deviation of the estimated coefficient
compute_abf <- function(z, s, prior_variance){
  abf <- sqrt(s^2/(s^2+prior_variance))*exp(z^2/2*(prior_variance/(s^2+prior_variance)))
  return(abf)
}


compute_approx_post_var <- function(z, s, prior_variance){
  post_var <- 1/(1/s^2 + 1/prior_variance)
  return(post_var)
}

# @param post_var: posterior variance
# @param s: standard deviation of the estimated coefficient
# @param bhat: estimated beta effect
compute_approx_post_mean <- function(post_var, s, bhat){
  mu <- post_var/(s^2)*bhat
  return(mu)
}

```

```{r}
dat = readRDS("./data/sim_dat_simple.rds")

library(survival)
# Modified Karl's code for intercept part
devtools::load_all("/Users/nicholeyang/Desktop/logisticsusie")
```

```{r}
surv_uni_fun <- function(x, y, o, prior_variance, estimate_intercept = 0, ...){
  fit <- coxph(y~ x + offset(o))
  bhat <- summary(fit)$coefficients[1, 1] # bhat = -alphahat
  sd <- summary(fit)$coefficients[1, 3]
  zscore <- bhat/sd
  bf <- compute_abf(zscore, sd, prior_variance)
  var <- compute_approx_post_var(zscore, sd, prior_variance)
  mu <- compute_approx_post_mean(var, sd, bhat)
  lbf <- log(bf)
  return(list(mu = mu, var=var, lbf=lbf, intercept=0))
}

fit_coxph <- ser_from_univariate(surv_uni_fun)
```



### Data 1: simulated from null model with X independent

$\log T_i =\beta_0+\epsilon_i$ and $\beta_0 = 1$.

```{r}
## Create  survival object. status == 2 is death
dat[[1]]$y <- with(dat[[1]], Surv(surT, status == 2))
# Fit cox ph. Cox ph model with select multiple significant predictors..
cox1 <- coxph(y ~ .-status - surT, data =  dat[[1]])
```

#### Run IBSS on data from null model with L=1.
```{r  }
p = 50
X = as.matrix(dat[[1]][, c(2:(p+1))])
y = dat[[1]]$y
```

```{r }
# IBSS of susie
set.seed(1)
fit1 <- ibss_from_ser(X, y, L = 1, prior_variance = 1., prior_weights = rep(1/p, p), tol = 1e-3, maxit = 500, estimate_intercept = TRUE, ser_function = fit_coxph)
```

```{r}
par(mfrow = c(1,2))
hist(fit1$alpha, breaks = 20)
hist(fit1$mu* fit1$alpha, breaks = 20)
```

#### Run IBSS on data from null model, L = 5.
```{r  }
# IBSS of susie
t1 <- proc.time()
fit2 <- ibss_from_ser(X, y, L = 5, prior_variance = 1., prior_weights = rep(1/p, p), tol = 1e-3, maxit = 500, estimate_intercept = TRUE, ser_function = fit_coxph)
t2 <- proc.time()
t(apply(fit2$alpha, 1, function(x) sort(x, decreasing = TRUE)))
```

```{r}
beta <- colSums(fit2$alpha * fit2$mu)
hist(beta, breaks = 20)
```


### Data 2: simulated from null model with highly correlated X.

$\log T_i =\beta_0+\epsilon_i$ and $\beta_0 = 1$.

```{r}
## Create  survival object. status == 2 is death
dat[[2]]$y <- with(dat[[2]], Surv(surT, status == 2))
# Fit cox ph. Cox ph model with select multiple significant predictors..
cox1 <- coxph(y ~ .-status - surT, data =  dat[[2]])
```

#### Run IBSS on data from null model with L=1.
```{r  }
p = 50
X = as.matrix(dat[[2]][, c(2:(p+1))])
y = dat[[2]]$y
```

```{r }
# IBSS of susie
set.seed(1)
fit1 <- ibss_from_ser(X, y, L = 1, prior_variance = 1., prior_weights = rep(1/p, p), tol = 1e-3, maxit = 500, estimate_intercept = TRUE, ser_function = fit_coxph)
```

```{r}
par(mfrow = c(1,2))
hist(fit1$alpha, breaks = 20)
hist(fit1$mu* fit1$alpha, breaks = 20)
```

#### Run IBSS on data from null model, L = 5.
```{r  }
# IBSS of susie
t1 <- proc.time()
fit2 <- ibss_from_ser(X, y, L = 5, prior_variance = 1., prior_weights = rep(1/p, p), tol = 1e-3, maxit = 500, estimate_intercept = TRUE, ser_function = fit_coxph)
t2 <- proc.time()
t(apply(fit2$alpha, 1, function(x) sort(x, decreasing = TRUE)))
```

```{r}
beta <- colSums(fit2$alpha * fit2$mu)
hist(beta, breaks = 20)
```



### Data 3: simulated from $\log T_i=\beta_0 + \beta_1x_{i1}$. Predictors are independent. 

$\beta_0 = 1, \beta_1 = 3$
```{r  }
dat[[3]]$y <- with(dat[[3]], Surv(surT, status == 2))
X = as.matrix(dat[[3]][, c(2:(p+1))])
y = dat[[3]]$y

# IBSS of susie
t1 <- proc.time()
fit3 <- ibss_from_ser(X, y, L = 5, prior_variance = 1., prior_weights = rep(1/p, p), tol = 1e-3, maxit = 500, estimate_intercept = TRUE, ser_function = fit_coxph)
t2 <- proc.time()
t2 - t1
```

```{r}
fit3$alpha
beta <- colSums(fit3$alpha * fit3$mu)
hist(beta, breaks = 20)
```


### Data 4: simulated from $\log T_i=\beta_0 + \beta_1x_{i1}$. Predictors have high correlation. 
```{r  }
dat[[4]]$y <- with(dat[[4]], Surv(surT, status == 2))
X = as.matrix(dat[[4]][, c(2:(p+1))])
y = dat[[4]]$y

# IBSS of susie
t1 <- proc.time()
fit4 <- ibss_from_ser(X, y, L = 5, prior_variance = 1., prior_weights = rep(1/p, p), tol = 1e-3, maxit = 500, estimate_intercept = TRUE, ser_function = fit_coxph)
t2 <- proc.time()
t2 - t1
t(apply(fit4$alpha, 1, function(x) sort(x, decreasing = TRUE)))
```

```{r}
beta <- colSums(fit4$alpha * fit4$mu)
hist(beta, breaks = 20)
```


### Data 5: simulated from $\log T_i=\beta_0 + \beta_1x_{i1} + \beta_1x_{i2}$. Predictors have high correlation

$\beta_0 = 1, \beta_1 = 3, \beta_2 = 1.5$ and $cor=0.9$.
```{r  }
dat[[5]]$y <- with(dat[[5]], Surv(surT, status == 2))
X = as.matrix(dat[[5]][, c(2:(p+1))])
y = dat[[5]]$y
 
# IBSS of susie
t1 <- proc.time()
fit5 <- ibss_from_ser(X, y, L = 5, prior_variance = 1., prior_weights = rep(1/p, p), tol = 1e-3, maxit = 500, estimate_intercept = TRUE, ser_function = fit_coxph)
t2 <- proc.time()
t2 - t1

```

```{r}
fit5$alpha
t(apply(fit5$alpha, 1, function(x) sort(x, decreasing = TRUE)))
```

```{r}
beta <- colSums(fit5$alpha * fit5$mu)
hist(beta, breaks = 20)
```

