---
title: "Run survival.svb on simple sim data"
author: "Yunqi Yang"
date: "3/3/2023"
output: html_document
---

## Description:

The candidate method is called **survival.svb**. 

Publication link: https://doi.org/10.1093/bioinformatics/btac416

Software link:https://github.com/mkomod/survival.svb

**Conclusion**: Scenario 2 doesn't work well. But all other scenarios work. 
```{r}
library(survival)
library(survival.svb)
```

### Example code from their package
```{r eval = FALSE}

# Question: Why re-scale censored data???
n <- 200                        # number of sample
p <- 1000                       # number of features
s <- 10                         # number of non-zero coefficients
censoring_lvl <- 0.4            # degree of censoring


# generate some test data
set.seed(1)
b <- sample(c(runif(s, -2, 2), rep(0, p-s)))
X <- matrix(rnorm(n * p), nrow=n)
Y <- log(1 - runif(n)) / -exp(X %*% b)
delta  <- runif(n) > censoring_lvl   		# 0: censored, 1: uncensored
Y[!delta] <- Y[!delta] * runif(sum(!delta))	# rescale censored data


# fit the model
f <- survival.svb::svb.fit(Y, delta, X)

# plot the results
plot(b, xlab=expression(beta), main="Coefficient value", pch=8, ylim=c(-2,2))
# effect size mu * gamma
points(f$m * f$g, pch=20, col=2)
legend("topleft", legend=c(expression(beta), expression(hat(beta))),
       pch=c(8, 20), col=c(1, 2))

```

```{r}
dat = readRDS("./data/sim_dat_simple.rds")
```

### Data 1: null model with X independent
```{r}
# In survival.svb, 0: censored & 1: uncensored
p = 50
X = as.matrix(dat[[1]][, c(2:(p+1))])
y = dat[[1]]$surT
delta = rep(1, length(dat[[1]]$status))
```

```{r}
f <- survival.svb::svb.fit(y, delta, X)
f$inclusion_prob
# effect size
f$beta_hat
```
```{r}
sort(f$inclusion_prob, decreasing = TRUE)
```


### Data 2: simulated from null model with highly correlated X. corr = 0.9

Wrongly selected two effects of different signs, cancelled out...
```{r}
# In survival.svb, 0: censored & 1: uncensored
p = 50
X = as.matrix(dat[[2]][, c(2:(p+1))])
y = dat[[2]]$surT
delta = rep(1, length(dat[[2]]$status))
```

```{r}
f <- survival.svb::svb.fit(y, delta, X)
f$inclusion_prob
# effect size
f$beta_hat
```
```{r}
sort(f$inclusion_prob, decreasing = TRUE)
```


### Data 3: simulated from one predictor model. Predictors are independent.
```{r}
# In survival.svb, 0: censored & 1: uncensored
p = 50
X = as.matrix(dat[[3]][, c(2:(p+1))])
y = dat[[3]]$surT
delta = rep(1, length(dat[[3]]$status))
```

```{r}
f <- survival.svb::svb.fit(y, delta, X)
sort(f$inclusion_prob, decreasing = TRUE)
```


### Data 4: simulated from one predictor model. Predictors are highly correlated, corr = 0.9
```{r}
# In survival.svb, 0: censored & 1: uncensored
p = 50
X = as.matrix(dat[[4]][, c(2:(p+1))])
y = dat[[4]]$surT
delta = rep(1, length(dat[[4]]$status))
```

```{r}
f <- survival.svb::svb.fit(y, delta, X)
sort(f$inclusion_prob, decreasing = TRUE)
f$inclusion_prob
```


### Data 5: simulated from two predictor model. Predictors have high correlation, corr = 0.9

```{r}
# In survival.svb, 0: censored & 1: uncensored
p = 50
X = as.matrix(dat[[5]][, c(2:(p+1))])
y = dat[[5]]$surT
delta = rep(1, length(dat[[5]]$status))
```

```{r}
f <- survival.svb::svb.fit(y, delta, X)
sort(f$inclusion_prob, decreasing = TRUE)
```

