---
title: "survival_example"
author: "Yunqi Yang"
date: "2/1/2023"
output: html_document
---

### Description: run a survival analysis example from:
https://rstudio-pubs-static.s3.amazonaws.com/5896_8f0fed2ccbbd42489276e554a05af87e.html


## Data description: NCCTG Lung Cancer Data

Description:
     Survival in patients with advanced lung cancer from the North
     Central Cancer Treatment Group.  Performance scores rate how well
     the patient can perform usual daily activities.

inst:       Institution code
time:       Survival time in days
status:     censoring status 1=censored, 2=dead
age:        Age in years
sex:        Male=1 Female=2
ph.ecog:    ECOG performance score (0=good 5=dead)
ph.karno:   Karnofsky performance score (bad=0-good=100) rated by physician
pat.karno:  Karnofsky performance score as rated by patient
meal.cal:   Calories consumed at meals
wt.loss:    Weight loss in last six months
```{r}
## Load survival package
library(survival)
## List datasets in survival package
# data(package = "survival")


## Load lung data
data(lung)
head(lung)
```

## Create a survival object
```{r}
## Add survival object. status == 2 is death
lung$SurvObj <- with(lung, Surv(time, status == 2))
## Check data
head(lung)
```

## Cox regression using coxph
```{r}
## Fit Cox regression: age, sex, Karnofsky performance score, wt loss
res.cox1 <- coxph(SurvObj ~ age + sex + ph.karno + wt.loss, data =  lung)
res.cox1
```

```{r}
## Check for violation of proportional hazard (constant HR over time)
(res.zph1 <- cox.zph(res.cox1))
```

```{r fig.width = 8, fig.height=8}
## Displays a graph of the scaled Schoenfeld residuals, along with a smooth curve.
par(mfrow = c(2,2))
plot(res.zph1)
```

