---
title: "Survival susie vs. survival susie rss"
author: "Yunqi Yang"
date: "05/08/2023"
output: html_document
---

### Description:

Compare results for susie and susierss on real genotype data from UKBiobank with $n=50000$ and $p=1000$. SuSIE and SuSIE-RSS run for 10 iterations. 

Would be interesting to compare with SuSiE(10 iterations) and SUSIE-RSS (100 or 500 iterations). 


```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE,results = "hold",
                      fig.align = "center")
```

```{r}

calculate_tpr_vs_fdr <- function(pip, is_effect, ts){
  res <- matrix(NA, nrow = length(ts), ncol = 2)
  colnames(res) = c("tpr", "fdr")
  for (i in 1:length(ts)){
    pred_pos = pip >= ts[i]
    tp = pip >= ts[i] & is_effect == 1
    fp = pip >= ts[i] & is_effect == 0
    tpr = sum(tp)/sum(is_effect)
    fdr = sum(fp)/sum(pred_pos)
    res[i, ] = c(tpr, fdr)
  }
  return(res)
}

# coverage: the proportion of CSs that contain an effect variable
# @param dat_indx: the indx for the data from dsc
# @param res.cs: credible sets from dsc
calculate_cs_coverage = function(res.cs, res.is_effect, dat_indx){
  contain_status = c()
  for (indx in dat_indx){
     cs = res.cs[[indx]]$cs
     true_effect = which(res.is_effect[[indx]] >= 1)
     if (!is.null(cs)){
       for (j in 1:length(cs)){
         res = ifelse(sum(true_effect %in% unlist(cs[j])) ==  1, 1, 0)
         contain_status = c(contain_status, res)
        }
     }
  }
  coverage = sum(contain_status)/length(contain_status)
  return(coverage)
}

# @param res.cs: credible sets from dsc
# @param dat_indx: the indx for the data from dsc
# @p: number of variables in each simulation replicate. 
get_cs_effect = function(res.cs, dat_indx, p){
  cs_effect = c()
  for (indx in dat_indx){
    effect = rep(0, p)
    cs_effect_indx = c(unlist(res.cs[[indx]]$cs))
    effect[cs_effect_indx] = 1
    cs_effect = c(cs_effect, effect)
  }
  return(cs_effect)
}
```

```{r}
susie = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/susie.rds")
rss = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/susie.rss.rds")
rss.varY = readRDS("/project2/mstephens/yunqiyang/surv-susie/dsc202401/susie.rss.varY.rds")
```

### 1. Results using real correlation structure from data

```{r}
t95 = 951
```

```{r, fig.width=12, fig.height=8}
par(mfrow = c(2,3), cex.axis = 1.5)
censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "real" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.rss = unlist(lapply(indx, function(x) rss$susie_rss.pip[[x]]))
  pip.rss.varY = unlist(lapply(indx, function(x) rss.varY$susie_rss_varY.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 0.99, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.rss = calculate_tpr_vs_fdr(pip.rss, is_effect, ts)
  res.rss.varY = calculate_tpr_vs_fdr(pip.rss.varY, is_effect, ts)
  
  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0,1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Real correlation, effect 0-3", ",censor=", censor_lvl[i]))
  lines(res.rss[,2], res.rss[,1], type = "l", col = 2)
  lines(res.rss.varY[,2], res.rss.varY[,1], type = "l", col = 3)
  
  points(res.susie[t95,2], res.susie[t95, 1])
  points(res.rss[t95,2], res.rss[t95, 1])
  points(res.rss.varY[t95,2], res.rss.varY[t95, 1])
  
  legend("topleft", legend = c("susie", "susie.rss.varY=1", "susie.rss.varY"), col = c(1,2,3), lty = 1)
}

```
The dots indicate PIP threshold = 0.95




### 2. Results using independent X, without data from null model. 

```{r, fig.width=12, fig.height=8}
par(mfrow = c(2,3),cex.axis = 1.5)
censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)

for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i] & susie$simulate.num_effect != 0)
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.rss = unlist(lapply(indx, function(x) rss$susie_rss.pip[[x]]))
  pip.rss.varY = unlist(lapply(indx, function(x) rss.varY$susie_rss_varY.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 0.99, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.rss = calculate_tpr_vs_fdr(pip.rss, is_effect, ts)
  res.rss.varY = calculate_tpr_vs_fdr(pip.rss.varY, is_effect, ts)

  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0,1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Independent X, effect 0-3", ",censor=", censor_lvl[i]))
  lines(res.rss[,2], res.rss[,1], type = "l", col = 2)
  lines(res.rss.varY[,2], res.rss.varY[,1], type = "l", col = 3)
  
  points(res.susie[t95,2], res.susie[t95, 1])
  points(res.rss[t95,2], res.rss[t95, 1])
  points(res.rss.varY[t95,2], res.rss.varY[t95, 1])
  
  legend("bottomright", legend = c("susie", "susie.rss.varY=1", "susie.rss.varY"), col = c(1,2,3), lty = 1)
  
}

```

The dots indicate PIP threshold = 0.95.


### 3. Results using independent X, with data from null model. 

```{r, fig.width=10, fig.height=8}
par(mfrow = c(2,3),cex.axis = 1.5)
censor_lvl = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)

for (i in 1:length(censor_lvl)){
  indx = which(susie$simulate.cor_type == "independent" & susie$simulate.censor_lvl == censor_lvl[i])
  pip.susie = unlist(lapply(indx, function(x) susie$susie.pip[[x]]))
  pip.rss = unlist(lapply(indx, function(x) rss$susie_rss.pip[[x]]))
  pip.rss.varY = unlist(lapply(indx, function(x) rss.varY$susie_rss_varY.pip[[x]]))
  is_effect = unlist(lapply(indx, function(x) susie$simulate.is_effect[[x]]))
  
  ts = seq(from = 0, to = 0.99, by = 0.001)
  res.susie = calculate_tpr_vs_fdr(pip.susie, is_effect, ts)
  res.rss = calculate_tpr_vs_fdr(pip.rss, is_effect, ts)
  res.rss.varY = calculate_tpr_vs_fdr(pip.rss.varY, is_effect, ts)

  plot(res.susie[,2], res.susie[,1], type = "l", xlim = c(0,1), ylim = c(0, 1), xlab = "FDR", ylab = "Power",
       main = paste0("Real correlation, effect 0-3", ",censor=", censor_lvl[i]))
  lines(res.rss[,2], res.rss[,1], type = "l", col = 2)
  lines(res.rss.varY[,2], res.rss.varY[,1], type = "l", col = 3)
  
  points(res.susie[t95,2], res.susie[t95, 1])
  points(res.rss[t95,2], res.rss[t95, 1])
  points(res.rss.varY[t95,2], res.rss.varY[t95, 1])
  
  legend("bottomright", legend = c("susie", "susie.rss.varY=1", "susie.rss.varY"), col = c(1,2,3), lty = 1)
}

```

The dots indicate PIP threshold = 0.95.


### 5. Assess Susie CS

```{r}
coverage = matrix(NA, ncol = 3, nrow = length(censor_lvl))
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
colnames(coverage) = c("effect:1", "effect:2", "effect:3")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:3){
  for (j in 1:6){
    dat_indx = which(susie$simulate.num_effect == i & susie$simulate.censor_lvl == censoring[j])
    coverage[j, i] = calculate_cs_coverage(susie$susie.cs, susie$simulate.is_effect, dat_indx)
  }
}

coverage
```

```{r}
power_cs = matrix(NA, ncol = 3, nrow = 6)
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
colnames(power_cs) = c("effect:1", "effect:2", "effect:3")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:3){
  for (j in 1:6){
    dat_indx = which(susie$simulate.num_effect == i & susie$simulate.censor_lvl == censoring[j])
    cs_effect = get_cs_effect(susie$susie.cs, dat_indx, p = 1000)
    is_effect = unlist(lapply(dat_indx, function(x) susie$simulate.is_effect[[x]]))
    power = sum(cs_effect ==1 & is_effect == 1)/sum(is_effect)
    power_cs[j, i] = power
  }
}

power_cs
```


```{r}
coverage = matrix(NA, ncol = 2, nrow = 6)
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
cor_type = c("real", "independent")
colnames(coverage) = c("real correlation", "independent")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:2){
  for (j in 1:6){
    dat_indx = which(susie$simulate.num_effect != 0 & susie$simulate.cor_type == cor_type[i] & susie$simulate.censor_lvl == censoring[j])
    coverage[j, i] = calculate_cs_coverage(susie$susie.cs, susie$simulate.is_effect, dat_indx)
  }
}

coverage
```

```{r}
power_cs = matrix(NA, ncol = 2, nrow = 6)
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
colnames(power_cs) = c("real correlation", "independent")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:2){
  for (j in 1:6){
    dat_indx = which(susie$simulate.num_effect != 0 & susie$simulate.cor_type == cor_type[i] & susie$simulate.censor_lvl == censoring[j])
    cs_effect = get_cs_effect(susie$susie.cs, dat_indx, p = 1000)
    is_effect = unlist(lapply(dat_indx, function(x) susie$simulate.is_effect[[x]]))
    power = sum(cs_effect ==1 & is_effect == 1)/sum(is_effect)
    power_cs[j, i] = power
  }
}

power_cs
```


### 6. Assess Susie.rss cs

```{r}
coverage = matrix(NA, ncol = 3, nrow = 6)
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
colnames(coverage) = c("effect:1", "effect:2", "effect:3")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:3){
  for (j in 1:6){
    dat_indx = which(rss$simulate.num_effect == i & rss$simulate.censor_lvl == censoring[j])
    coverage[j, i] = calculate_cs_coverage(rss$susie_rss.cs, rss$simulate.is_effect, dat_indx)
  }
}

coverage
```

```{r}
power_cs = matrix(NA, ncol = 3, nrow = 6)
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
colnames(power_cs) = c("effect:1", "effect:2", "effect:3")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:3){
  for (j in 1:6){
    dat_indx = which(rss$simulate.num_effect == i & rss$simulate.censor_lvl == censoring[j])
    cs_effect = get_cs_effect(rss$susie_rss.cs, dat_indx, p = 1000)
    is_effect = unlist(lapply(dat_indx, function(x) rss$simulate.is_effect[[x]]))
    power = sum(cs_effect ==1 & is_effect == 1)/sum(is_effect)
    power_cs[j, i] = power
  }
}

power_cs
```


```{r}
coverage = matrix(NA, ncol = 2, nrow = 6)
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
cor_type = c("real", "independent")
colnames(coverage) = c("real correlation", "independent")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:2){
  for (j in 1:6){
    dat_indx = which(rss$simulate.num_effect != 0 & rss$simulate.cor_type == cor_type[i] & rss$simulate.censor_lvl == censoring[j])
    coverage[j, i] = calculate_cs_coverage(rss$susie_rss.cs, rss$simulate.is_effect, dat_indx)
  }
}

coverage
```

```{r}
power_cs = matrix(NA, ncol = 2, nrow = 6)
censoring = c(0, 0.2, 0.4, 0.6, 0.8, 0.99)
colnames(power_cs) = c("real correlation", "independent")
rownames(coverage) = c("censor:0", "censor:0.2", "censor:0.4", "censor:0.6", "censor:0.8", "censor:0.99")
for (i in 1:2){
  for (j in 1:6){
    dat_indx = which(rss$simulate.num_effect != 0 & rss$simulate.cor_type == cor_type[i] & rss$simulate.censor_lvl == censoring[j])
    cs_effect = get_cs_effect(rss$susie_rss.cs, dat_indx, p = 1000)
    is_effect = unlist(lapply(dat_indx, function(x) rss$simulate.is_effect[[x]]))
    power = sum(cs_effect ==1 & is_effect == 1)/sum(is_effect)
    power_cs[j, i] = power
  }
}

power_cs
```

